---
title: 'Data Lab 10: Network Visualizations'
author: "Anin Nair"
date: "``13-11-2023"
output:
  html_document:
    df_print: paged
---

### 1. Setting up the libraries and datsets
Loading all the libraries needed and importing the CO2 Emissions  CSV file using the *read.csv* function and then loading the first few lines of data using the *head* function.

```{r}
library (tidyverse)
library(ggplot2)
library(ggrepel)
library(ggnetwork)
library(network)
nodes <- read.csv("artists_2012_2020.csv")
edges <- read.csv("collaboration_edgelist_2012_2020.csv")
head(edges)
head(nodes)
```

### 2.Setting up the nodes and the edges
```{r}
knitr::kable(nodes)
knitr::kable(edges)
network_data <- network(x=edges, vertices=nodes,directed=FALSE)
network_data
```
### 3. Network Plot - 1
Here I've plotted a network visualization of artists and Grammy nominees. This first plot specifies the gender of the artists. There is no specific reason as to why I chose the color, I just felt blue and violetred where suited well for a binary variable and also color-blind friendly. I believe that none of the layouts did good enough for this particular dataset but I chose this one because the _fruchtermanreingold_ is the most visually appealing. 

```{r}
net <- ggnetwork(network_data,
layout="fruchtermanreingold")
ggplot(net, aes(x=x,y=y,xend=xend,yend=yend)) + geom_edges() +geom_nodes(aes(color = as.factor(Gender)), size = 2.5) + labs(x = " ", y = " " , color = "Gender")+   ggtitle("Network Visualization of Grammy Nominees by Gender ") + scale_color_manual(labels = c("Male", "Female"), values = c("blue", "violetred"))
#+ geom_nodelabel_repel(aes(label = nodes$Artist)) + theme_blank()

```
## 4. Network Plot - 2
Here I've plotted a network visualization of artists and Grammy nominees. This first plot specifies if they won a Grammy or not. There is no specific reason as to why I chose the color, I just felt felt blue and violetred suited well for a binary variable and also color-blind friendly. I believe that none of the layouts did good enough for this particular dataset but I chose this one because the _fruchtermanreingold_ is the most visually appealing. 
```{r}
net <- ggnetwork(network_data,
layout="fruchtermanreingold")
ggplot(net, aes(x=x,y=y,xend=xend,yend=yend)) + geom_edges() +geom_nodes(aes(color = as.factor(GrammyWinner)), size = 2.5) + labs(x = " ", y = " " , color = "Grammy Winners")+ ggtitle("Network Visualization of Grammy Nominees by Winners ")+ scale_color_manual(labels = c("Not a Winner", "Winner"), values = c("blue", "violetred"))
```
## Note

I was unfortunately unable to add labels to both of the plots. I tried all possible ways but I kept getting an error "Artist not found" even though I have a variable named Artist. I can email to you the image, because I'm not sure how to attach it on an R notebook file!
